### 2008-09-17 trunk/Papers/rmrs-08/conclusion.tex
iterated over whole paper

\begin{abstract}
  la la la
\end{abstract}

>>>>

\section{Conclusion}
\label{sec:conclusion}

In this paper, we have motivated and defined \rmrs---a semantic
framework that has been used to represent, compare, and combine
semantic information computed from deep and shallow parsers.  \rmrs\
is designed to be maximally flexible on the type of semantic
information that can be left underspecified -- it can underspecify
lexical ambiguity, predicate-argument relations, and scope --, so that
the semantic output of a shallow parser needn't over-determine or
under-determine the semantic information that can be extracted from
the shallow syntactic analysis.  This lays the formal foundations for
a semantic representation formalism that is emerging as a standard in
robust semantic processing.

We then defined the concept of \emph{solved forms} of an \rmrs, in
analogy to dominance graphs.  Solved forms abstract away from
irrelevant differences between models, and have been a crucial concept
in the development of efficient solvers for dominance graphs.  In the
future, we hope to draw upon that body of literature to develop
efficient algorithms for processing \rmrs s, including solvers which
will compute fully specific semantic representations from an \rmrs,
efficient inference algorithms, and algorithms for entailment
checking.  We have argued that entailment is a central notion in
comparing the outputs of different NLP tools, and provided a syntactic
characterization for \rmrs\ entailment, which we hope will allow us to
develop efficient entailment algorithms in the future.

### 2008-09-17 trunk/Papers/rmrs-08/entailment.tex
iterated over whole paper

Now that we have the formal tools in place, let's go back to the
problem of checking compatibility of semantic outputs.  We can phrase
this as RMRS entailment.  This is progress already.

Then show the representation theorem of entailment as solved form
comparison:

\begin{thm}
  Let $\varphi, \varphi'$ be two RMRSs.  Then $\varphi \models
  \varphi'$ iff for every solved form $S$ of $\varphi$, there is a
  solved form $S'$ of $\varphi'$ such that $S$ is an extension of
  $S'$.
\end{thm}

(Of course, this doesn't give us an efficient algorithm for entailment
checking, because the number of solved forms may be exponential on
each side.  And we still don't know the complexity of deciding
entailment (even for normal dominance constraints).  Nevertheless,
this is nice, because it illustrates that our model theory is both
pretty and useful for proving stuff.)

>>>>

The definitions in the previous section serve an important theoretical
purpose in that they provide a formal underpinning for \rmrs, which so
far has been used for partial semantic representations in implemented
systems despite the lack of a formal definition.  Next to the peace of
mind that comes with the use of a formally understood formalism, a
clean definition can also provide a foundation for the development of
efficient algorithms; this has been shown in the past in the dominance
graph literature, most recently \newcite{KolRegTha08}.  We hope that
this paper can likewise provide a starting point for the development
of efficient solvers and inference mechanisms for \rmrs.

On the other hand, the above definition of \rmrs\ provides new
concepts which can help us phrase questions of practical grammar
engineering in well-defined formal terms.  For instance, one crucial
issue in developing a hybrid system that combines the outputs of deep
and shallow processors is to determine whether the \rmrs s produced by
the two systems are compatible.  Because we assume that the deep
grammar is more informative than that of the shallow processor, this
amounts to checking whether every piece of information in the shallow
representation is consistent with what the deep grammar said.  In
general, it isn't even obvious what ``consistency'' is supposed to
mean exactly, given that variables might be renamed, $\ARG$ roles
assigned in different ways, and structural relations underspecified.
But given the formal definition of \rmrs, we can now propose that an
\rmrs\ $\varphi'$ that was computed by a shallow processor is
consistent with an \rmrs\ $\varphi$ that was computed by a deep
grammar amounts to checking \emph{entailment} between the two \rmrs s,
i.e.\ whether $\varphi \models \varphi'$.  If we could establish
efficiently that $\varphi'$ is entailed by $\varphi$, this would tell
us that all models that satisfy $\varphi$ are also models of
$\varphi'$, i.e.\ the deep representation is formally more specific.

We cannot provide an efficient algorithm for testing entailment in
this paper.  However, we propose the following novel syntactic
characterization of entailment.

\begin{definition}
  We call an \rmrs\ $\varphi'$ an \emph{extension} of an \rmrs\
  $\varphi$ iff \todo{fill this in -- don't have the def with me right
    now}.
\end{definition}

\begin{thm}\label{thm:big-one}
  Let $\varphi, \varphi'$ be two {\sc rmrs}s.  Then $\varphi \models
  \varphi'$ iff for every solved form $S$ of $\varphi$, there is a
  solved form $S'$ of $\varphi'$ such that $S$ is an extension of
  $S'$. \todo{We don't know if this is true for qeq; the proof is
    written for $\dom$.}  \todo{Also note that this is about a version
    of entailment in which $\varphi$ and $\varphi'$ contain the same
    variables, which makes the claim a bit weaker.}
\end{thm}
\begin{proof}[Proof (sketch)]
``$\Leftarrow$'' follows from
Props.~\ref{prop:solved-forms-are-satisfiable} and
\ref{prop:models-satisfy-solved-forms}.

``$\Rightarrow$'': The basic idea is to construct a solved form for
$\varphi'$ by choosing a solved form for $\varphi$ and removing all
atoms using variables that don't occur in $\varphi'$.  The hard part
is the proof that the result is a solved form of $\varphi'$; this step
involves proving that if $\varphi \models \varphi'$, then all EPs in
$\varphi'$ also occur in $\varphi$.  
\end{proof}

This characterization does not by itself provide an efficient test for
entailment because the number of solved forms of $\varphi$ and
$\varphi'$ may be exponential, and thus it is infeasible to compute
them all and then test for extensions.  Nevertheless, we hope that the
theorem provides a first step towards an improved understanding of
entailment that will lead to such an efficient algorithm in the
future.

### 2008-09-17 trunk/Papers/rmrs-08/introduction.tex
iterated over whole paper

On the other hand, there has been a recent trend towards systems for
robust wide-coverage semantic construction.  For instance,
\cite{bos:etal:2004} exploit
{\sc ccg}'s elegant syntax-semantics interface to produce a unique
logical form for each derivation of
a statistical {\sc ccg} parser;
\cite{zettlemoyer:collins:2007} and others learn
string-to-logical-form mappings, exploting informative features from
shallow language processors; and \cite{copestake:2003} and
\cite{frank:2004}
endow robust language processors with semantic components.

However, there are certain semantic
phenomena that these robust
approaches don't capture reliably, including quantifier scope, optional
arguments,
% COMMENT: this issue about optional arguments is really too cryptic.
% Basically, there is no way for assessing by *any* learned language
% processor whether a single word form that appears in
% a number of different syntactic valencies is an example of lexical
% sense ambiguity, or they are all examples of the same sense being
% used in syntactically different ways (i.e., we have optional
% arguments).  Only hand-crafted grammars can tell you this.  So, for
% example, Bos et al (2004) would get bet(x,y) for "I bet 5 pounds",
% and bet(x,y,z) for "I bet you 5 pounds" and bet(x,p) for "I bet it's
% going to rain": I.e., the vocab of the semantic formalism is
% constructed automatically from NL word forms, and the predicate
% symbols in that automatically constructed vocab vary in (a) valency,
% and (b) the semantic dependencies represented by a given argument
% position (e.g., the second argument of "bet" in "I bet 5 pounds" is
% 5 pounds, and for "I bet it's raining" it's "it's raining".
% 
and long-distance dependencies (for instance, \cite{clark:etal:2004}
report that the parser used in \cite{bos:etal:2004} yields 63\%
accuracy on object
extraction; e.g., {\em the man that I met\dots}).
Forcing a shallow parser to make a decision about these
phenomena can therefore be error-prone.  Depending on the
application, it may be preferrable to give the parser the option to to
leave the decision open when it's not sufficiently informed---i.e., to
compute a partial semantic representation that neither over-determines
nor under-determines the semantic information revealed by the
(shallow) syntactic analysis, and to complete the semantic information
later, using information extraneous to the
parser.

In this paper, we focus on one particular approach to representing
partial semantic information:
Robust Minimal Recursion Semantics ({\sc rmrs},
\cite{copestake:2003}).  {\sc rmrs} is a generalisation of {\sc mrs}
\cite{copestake:etal:2005} that is designed to support underspecification
of scope and predicate-argument structure.
\cite{copestake:2003,frank:2004} use it to
specify semantic components to shallow parsers
ranging in depth from {\sc pos} taggers to chunk parsers  
and intermediate parsers such as {\sc rasp}
\cite{briscoe:etal:2006}.  {\sc mrs} analyses
derived from deep grammars, such as the English Resource Grammar
({\sc erg}, \cite{copestake:flickinger:2000}) are
special cases of 
{\sc rmrs}.

The key contribution we make is to cast {\sc rmrs}, for the first
time, as a logic with a clearly defined model theory.  This has a
number of advantages.  First, it allows us to formally relate \rmrs\ to
other frameworks such as \mrs\ and predicate logic in order to
establish the compatibility of \rmrs\ with the approaches to robust
semantic analysis mentioned above.  Second, it provides us with a
notion of logical entailment, which we can use to verify the
compatibility of \rmrs\ representations computed from different
parsers, such as a deep and a shallow parser.  Finally, it is only
through rigid formalisation that we can give users of \rmrs\ access to
inference systems and to algorithms for efficiently computing the
possible fully specific semantic representations described by the \rmrs\
output of a shallow parser.

>>>>

On the other hand, there has been a recent trend towards systems for
robust wide-coverage semantic construction.  For instance,
\newcite{bos:etal:2004} exploit {\sc ccg}'s elegant syntax-semantics
interface to produce a unique logical form for each derivation of a
statistical {\sc ccg} parser; \newcite{zettlemoyer:collins:2007} and
others learn string-to-logical-form mappings, exploting informative
features from shallow language processors; and
\newcite{copestake:2003} and \newcite{frank:2004} endow robust
language processors with semantic components.

% COMMENT: this issue about optional arguments is really too cryptic.
% Basically, there is no way for assessing by *any* learned language
% processor whether a single word form that appears in
% a number of different syntactic valencies is an example of lexical
% sense ambiguity, or they are all examples of the same sense being
% used in syntactically different ways (i.e., we have optional
% arguments).  Only hand-crafted grammars can tell you this.  So, for
% example, Bos et al (2004) would get bet(x,y) for "I bet 5 pounds",
% and bet(x,y,z) for "I bet you 5 pounds" and bet(x,p) for "I bet it's
% going to rain": I.e., the vocab of the semantic formalism is
% constructed automatically from NL word forms, and the predicate
% symbols in that automatically constructed vocab vary in (a) valency,
% and (b) the semantic dependencies represented by a given argument
% position (e.g., the second argument of "bet" in "I bet 5 pounds" is
% 5 pounds, and for "I bet it's raining" it's "it's raining".
%

However, there are certain semantic phenomena that these robust
approaches don't capture reliably, including quantifier scope,
optional arguments, and long-distance dependencies (for instance,
\newcite{clark:etal:2004} report that the parser used by
\newcite{bos:etal:2004} yields only 63\% accuracy on object
extraction; e.g., {\em the man that I met\dots}).  Forcing a shallow
parser to make a decision about these phenomena can therefore be
error-prone.  Depending on the application, it may be preferrable to
give the parser the option to to leave the decision open when it's not
sufficiently informed---i.e., to compute a partial semantic
representation that neither over-determines nor under-determines the
semantic information revealed by the (shallow) syntactic analysis, and
to complete the semantic information later, using information
extraneous to the parser.

In this paper, we focus on one particular approach to representing
partial semantic information: Robust Minimal Recursion Semantics
(\rmrs, \cite{copestake:2003}).  \rmrs\ is a generalisation of \mrs\
\cite{copestake:etal:2005} that is designed to support
underspecification of scope and predicate-argument structure.  It is
an emerging standard representation language for partial semantic
information, and has been applied in several implemented systems.  For
instance, \newcite{copestake:2003} and \newcite{frank:2004} use it to
specify semantic components to shallow parsers ranging in depth from
{\sc pos} taggers to chunk parsers and intermediate parsers such as
{\sc rasp} \cite{briscoe:etal:2006}.  {\sc mrs} analyses derived from
deep grammars, such as the English Resource Grammar ({\sc erg},
\cite{copestake:flickinger:2000}) can be seen as special cases of {\sc
  rmrs}.

However, from a formal perspective, \rmrs\ is somewhat ad-hoc in that
it has no clearly defined model theory; i.e., it is not defined what,
exactly, any given \rmrs\ representation actually \emph{means}.  The
key contribution we make is to cast {\sc rmrs}, for the first time, as
a logic with a well-defined model theory.  This has a number of
advantages.  First, it allows us to formally relate \rmrs\ to other
frameworks such as \mrs\ and predicate logic in order to establish the
compatibility of \rmrs\ with the approaches to robust semantic
analysis mentioned above.  \todo{except we don't do this here --
  perhaps we should} Second, it provides us with a notion of logical
entailment, which we can use to verify the compatibility of \rmrs\
representations computed from different parsers, such as a deep and a
shallow parser.  Finally, it is only through rigid formalisation that
we can give users of \rmrs\ access to inference systems and to
algorithms for efficiently computing the possible fully specific
semantic representations described by the \rmrs\ output of a shallow
parser.

### 2008-09-17 trunk/Papers/rmrs-08/motivation.tex
iterated over whole paper

% Motivate what we're doing and provide background.  Example outputs of
% deep and shallow parsers.  Illustrate and state the problem.
%
% Try to find an eg where the comparison isn't easy to see (from Anna
% Ritchie's deliverable?). 
% NOTE: on Ritichie: Perhaps "Three dogs bark on Tuesday in
% June" from page 11?  It's hard to follow what's going on, from the
% sheer size of the RMRSs and the need to allow for variation on the
% variable names.  But I don't know how to extract this figure
% techinically; nor whether we have room.

We motivate and describe {\sc rmrs} with a toy
example.  Sentence 
(\ref{eg:cat}a) features a semantic scope ambiguity that, according to
most deep grammars, syntax
doesn't resolve.    The {\sc mrs} (\ref{eg:cat}b) derived by the
{\sc erg} \cite{copestake:flickinger:2000}
reflects this: (i) each predication is
labelled ($l_1$, $l_4$, etc); (ii) scopal arguments are indicated by
holes (e.g., $h_2$, $h_3$); (iii)
constraints on scope are expressed by the qeq conditions (roughly, $h
=_q l$ means that only zero or more quantifiers lie between the scopal
argument $h$ and the predication labelled by $l$ in any fully-resolved
logical form); and (iv) these scope
constraints allow for more than one way of `plugging' the holes with
labels.\footnote{The event variable in $\_\mbox{{\em fat}}\_j\_1$ is 
  for copulas. {\em The
    cat is fat} has logical form 
  $\mbox{\_the\_q\_1}(x,\mbox{\_cat\_n\_1}(x),\mbox{\_fat\_j\_1}(e,x))$.}
\begin{examples}
\item   \label{eg:cat}
\begin{subexamples}
\item
Every fat cat chased some dog.
% \item
% $\mbox{\_every\_q\_1}(x, \mbox{\_fat\_j\_1}(e',x) \wedge
% \mbox{\_cat\_n\_1}(x),$\\
% \hspace*{0.5in}$\mbox{\_some\_q\_1}(y, \mbox{\_dog\_n\_1}(y),
% \mbox{\_chase\_v\_1}(e_{spast},x) )$\\
% $\mbox{\_some\_q\_1}(y, \mbox{\_dog\_n\_1}(y),$\\
% \hspace*{0.5in}$\mbox{\_every\_q\_1}(x, \mbox{\_fat\_j\_1}(e',x)
% \wedge 
% \mbox{\_cat\_n\_1}(x), 
% \mbox{\_chase\_v\_1}(e_{spast},x)))$
\item   $l_1\handel\mbox{\_every\_q\_1}(x,h_2,h_3)$\\
$l_4\handel\mbox{\_fat\_j\_1}(e',x)$\\
$l_4\handel\mbox{\_cat\_n\_1}(x)$\\
$l_5\handel\mbox{\_chased\_n\_1}(e,x,y)$\\
$l_6\handel\mbox{\_some\_n\_1}(y,h_7,h_8)$\\
$l_9\handel\mbox{\_dog\_n\_1}(y)$\\
$h_2 =_q l_4, h_7 =_q l_9$
\end{subexamples}
\end{examples}
(\ref{eg:cat}b) also exhibits {\sc mrs}' compact description of
conjunction through label sharing (see $l_4$), and the {\sc
  erg}'s naming convention for predicate symbols: each lexical predicate is
constructed from the word lemma, its syntactic category and a sense
number, and the leading underscore distinguishes lexical predicates
from non-lexical ones.

Conceptually, this {\sc mrs} is a partial description of all its
possible (fully specific) interpretations in context.  
These are expressed in a so-called {\em base language}, where each
predicate symbol in the {\sc mrs} is a proxy for one or more
base-language constructors of the same arity.
Assuming a very close
correspondence between the {\sc mrs} predicate symbols and the
base-language constructors they 
stand for, the {\sc mrs}
(\ref{eg:cat}b) (partially) describes both the base-language formulae in
Figure~\ref{fig:1}, shown as trees to emphasise that an {\sc mrs} is a
partial description of the formulae's {\em form}; not its
(base-language) entailments. 
In essence, the labels in the {\sc mrs} reify the
nodes in the tree, and each of these trees encodes one way of equating
holes with labels that respects qeq constraints and binding conditions
on the variables.

\begin{figure}[h]

\includegraphics[width=6cm]{pic-cat-chased-dog}

Put trees in here

\caption{The Form of alternative base logical forms of
  (\ref{eg:cat}a).}
\label{fig:1}
\end{figure}

The {\sc mrs} (\ref{eg:cat}b) is a notational variant of the {\sc
  rmrs} (\ref{eg:cat}c). {\sc rmrs}s offer a more factorised
representation: each predicate symbol is unary and the other
arguments to its corresonding base-language constructor
are represented by separate binary relations on the unique {\em
  anchor} of the predicate symbol ($a_1, a_{41},\ldots$) together with
variable equalities (e.g., $x=x_1$).
\begin{examples}
\item   [\ref{eg:cat}]
\begin{subexamples}
\item   [c]
$l_1:a_1\handel\mbox{\_every\_q\_1}(x_1),$\\
\hspace*{0.1in}$\mbox{RSTR}(a_1,h_2),
\mbox{BODY}(a_1,h_3)$\\ 
$l_{41}:a_{41}\handel\mbox{\_fat\_j\_1}(e'), \mbox{ARG1}(a_{41},x_2)$\\
$l_{42}:a_{42}\handel\mbox{\_cat\_n\_1}(x_3)$\\
$l_5:a_5\handel\mbox{\_chase\_v\_1}(e_{spast})$,\\
\hspace*{0.1in}$\mbox{ARG1}(a_5,x_4),
\mbox{ARG2}(a_5,x_5)$\\ 
$l_6:a_6\handel\mbox{\_some\_q\_1}(x_6)$,\\
\hspace*{0.1in}$\mbox{RSTR}(a_6,h_7),
\mbox{BODY}(a_6,h_8)$\\ 
$l_9:a_9\handel\mbox{\_dog\_n\_1}(x_7)$\\
$h_2=_q l_{42}, l_{41}=l_{42}, h_7 =_q l_9$\\
$x_1=x_2, x_2=x_3, x_3=x_4,$\\
$x_5=x_6, x_5=x_7$
\end{subexamples}
\end{examples}

This factored representation allows one to build semantic components
to shallow parsers where lexical or syntactic information is absent.
An extreme example would be a {\sc 
  pos} tagger: one can build its semantic component simply by deriving
lexical predicate symbols from the word lemmas and tags and assigning each
predication a unique label, anchor and
argument, as in (\ref{eg:cat-pos}):
\begin{examples}
\item   \label{eg:cat-pos}
\begin{subexamples}
\item   Every\_{\sc at1} fat\_{\sc jj} cat\_{\sc nn1} chased\_{\sc
    vvd} some\_{\sc dd} dog\_{\sc nn1}
\item   $l_1:a_1:\mbox{\_every\_q}(x)$, \\
$l_{41}:a_{41}:\mbox{\_fat\_a}(e')$,\\
$l_{42}:a_{42}:\mbox{\_cat\_n}(x_3)$\\
$l_5:a_5:\mbox{\_chased\_v}(e_{past})$, \\
$l_6:a_6:\mbox{\_some\_q}(x_4)$, \\
$l_9:a_9:\mbox{\_dog\_n}(x_5)$
\end{subexamples}
\end{examples}

Semantic relations, sense tags and the arity of the base-language
constructors are
missing from (\ref{eg:cat-pos}b) because the {\sc pos} tagger doesn't
reveal information about syntactic constituency, word sense or lexical
subcategorisation respectively.  
But the {\sc rmrs}s (\ref{eg:cat}c) and
(\ref{eg:cat-pos}b) are entirely compatible, the former being more
specific than the latter.  At least, this holds so long as the
base-language constructors that the {\sc rmrs} predicate symbol
{\_name\_t\_s} corresponds to is a 
subset of those corresponding to the predicate symbol {\em \_name\_t},
for any word lemma {\em name}, {\sc pos} tag $t$ and sense $s$ (written 
$\mbox{{\em \_name\_t\_s}} \sqsubset \mbox{{\em name\_t}}$).

The values of non-scopal arguments are underspecified through the
absence of ARG relations and/or through the absence of
equalities and inequalities among the variables $x_1, x_2,\ldots$.
The need to
underspecify such semantic dependencies is required when
the shallow parser doesn't record syntactic dependencies.
Indeed, there is no single {\sc mrs} which is satisfied by both the
trees in Figure~\ref{fig:2}, assuming of course that disjunction is not part of
the {\sc mrs} language.  In words, these trees collectively specify
the (partial) semantic information that $x$ is either the second
argument (i.e., the direct object) or the third argument (i.e., the
indirect object) to {\em give}, but we don't know which.  
A shallow
processor that misses long distance dependencies in
(\ref{eg:which-dog}) cannot
discern that the bone is the second argument to {\em
  give} in (\ref{eg:which-dog}a), while the dog is the third argument
in (\ref{eg:which-dog}b).
\begin{examples}
\item   \label{eg:which-dog}
\begin{subexamples}
\item   To which dog did Kim give the bone?
\item   Which bone did Kim give the dog?
\end{subexamples}
\end{examples}
Thus to maximise the semantic contribution
of the parser while not overdetermining it, the phrases
{\em give the bone} vs.\
{\em give the dog} should whether the semantic index of the noun
phrase is the second or third argument to {\em give},
as shown in Figure~\ref{fig:2}.   This will be expressed in
{\sc rmrs} via the binary relation  $\mbox{{\em ARG}}_{\{2,3\}}(a,x)$
between the anchor $a$ of the predicate symbol for {\em give} and the
individual $x$.   More generally, $\mbox{{\em ARG}}_n(a,x)$ means that
$x$ is an argument to the predicate anchored at $a$, but its argument
position is unspecified.

\begin{figure}

\leaf{$e$}
\leaf{$\ldots$}
\leaf{$x$}
\leaf{$\ldots$}
\branch{4}{\_give\_v\_b1}
\tree
\hfill
\leaf{$e$}
\leaf{$\ldots$}
\leaf{$\ldots$}
\leaf{$x$}
\branch{4}{\_give\_v\_b1}
\tree

\caption{$x$ is a participant in a
  giving event, but $x$'s role is unknown.}
\label{fig:2}
\end{figure}

The binary ARG relations also supports underspecifying the {\em arity}
of the base-language constructors, as is required when lexical
subcategorisation information is missing from the shallow parser.
While a deep parser assumes complete information about lexical
subcategorisation in any of its derivations, thereby fixing the arity
of the base-language constructors in the fully-specific logical form
to be the same as the {\sc mrs} predicate symbol $P$ that's introduced by
the lexeme, the predicate symbols in an {\sc rmrs} yielded from a
shallow parse should, in principle at least,
correspond to base language constructors of mixed arity.  For
instance, the English verb {\em bake} can be used causatively ({\em
  Kim baked the potatoes}), in which case semantically it corresponds
to a 3-place predicate symbol {\em \_bake\_v\_b1} in the base
language, or it may be used 
inchoatively ({\em the potatoes are baking}), that semantically
corresponds to a 2-place predicate {\em \_bake\_v\_b2}.  So the 
predicate symbol {\em 
  \_bake\_v} in the {\sc rmrs} that's introduced by a {\sc pos} tagger
must be able to denote either of these
base-language predicates in the {\sc rmrs}'s interpretation.

\hidden{
Finally, {\sc rmrs} allows one to underspecify the {\em sort} of an
argument by including in its language meta-variables $u_1,u_2,\ldots$
over individual, event, and handle variables.  For instance, the predicate
symbol {\em \_be\_v} produced by a {\sc pos} tagger is resolvable to several
3-place base-language constructors whose argument-types 
vary: the base-language constructor corresponding
to its use in {\em John is a man} takes an event and two individuals
as arguments; but in {\em It is John who talked} the constructor takes
an event, an individual and a proposition as arguments.
}

>>>>

We start with an informal discussion of some toy \rmrs s, which are
meant to illustrate how \rmrs\ can be used to represent partial
semantic information, and why people find it useful for this purpose.

\newcommand{\sem}[1]{\mathsf{#1}}
\newcommand{\sempred}[1]{\mathrm{#1}}

Consider the following sentence:
\begin{examples}
  \item Every fat cat chased some dog.
\end{examples}

This sentence exhibits several kinds of ambiguity, including a scope
ambiguity between ``every fat cat'' and ``some dog'' and lexical
ambiguities of the words ``cat'' and ``dog'' (which have 8 and 7 noun
senses according to Wordnet, respectively).  Two of these readings are
shown as logical forms below; they can be represented as trees as
shown in Fig.~\ref{fig:1}.

\begin{examples}
\item $\sem{\_every\_q\_1}(x, \sem{\_fat\_j\_1}(e',x) \wedge
    \sem{\_cat\_n\_1}(x), \sem{\_some\_q\_1}(y, \sem{\_dog\_n\_1}(y),
    \sem{\_chased\_v\_1}(e,x,y)))$
\label{ex:fat-cat-1}
\item $\sem{\_some\_q\_1}(y, \sem{\_dog\_n\_2}(y),
\sem{\_every\_q\_1}(x, \sem{\_fat\_j\_1}(e',x) \wedge
    \sem{\_cat\_n\_2}(x), 
    \sem{\_chased\_v\_1}(e,x,y)))$
\label{ex:fat-cat-2}
\end{examples}

\begin{figure*}[t]
\includegraphics[width=6cm]{pic-cat-chased-dog}
\hspace{1cm}
\includegraphics[width=6cm]{pic-cat-chased-dog-2}
\caption{Structure trees of the semantic representations (\ref{ex:fat-cat-1}) and
  (\ref{ex:fat-cat-2}). \label{fig:1}}
\end{figure*}

Now imagine that we are trying to extract semantic information from
the output of a part-of-speech tagger by using the word lemmas as
lexical predicate symbols.  Such a semantic representation is highly
partial, as it will not say anything about predicate-argument
relations, or resolve lexical ambiguity -- it will use predicate
symbols such as $\sempred{\_cat\_n}$, which might resolve to the
predicate symbols $\sem{\_cat\_n\_1}$ or $\sem{\_cat\_n\_2}$ in the
complete semantic representation.  (Notice that we are using different
fonts for the ambiguous and unambiguous predicate symbols.)  We can
express this information in \rmrs\ as follows.

\begin{examples}
\item \label{ex:cat-pos}
$l_1:a_1:\sempred{\_every\_q}(x)$, \\
$l_{41}:a_{41}:\sempred{\_fat\_a}(e')$,\\
$l_{42}:a_{42}:\sempred{\_cat\_n}(x_3)$\\
$l_5:a_5:\sempred{\_chased\_v}(e_{past})$, \\
$l_6:a_6:\sempred{\_some\_q}(x_4)$, \\
$l_9:a_9:\sempred{\_dog\_n}(x_5)$
\end{examples}

This \rmrs\ consists of six \emph{elementary predications} (EPs), one
for each word lemma in the sentence; each of them is prefixed by a
\emph{label} and an \emph{anchor}, which are essentially variables
that refer to nodes in the trees in Fig.~\ref{fig:1}.  We can say that
the two trees \emph{satisfy} this \rmrs\ because it is possible to map
the labels and anchors into each tree in such a way that the label of
the node they are mapped to is consistent with the label specified in
the elementary predication.  For instance, the first tree satisfies
the \rmrs\ because we can map $l_1$ and $a_1$ to the root of that
tree, and the root label $\sem{\_every\_q\_1}$ is consistent with the
EP predicat $\sempred{\_every\_q}$.

There are of course many other trees (and thus, fully specific
semantic representations such as (\ref{ex:fat-cat-1})) that are
described equally well by the same \rmrs; this is not surprising,
given that the semantic information we got from the POS tagger is so
incomplete.  If we have more information, say information about the
predicate-argument structure and semantic dependencies from a partial
parser such as \todo{what? CASS?}, we can represent these in a more
detailed \rmrs, as follows:

\begin{examples}
\item 
$l_1:a_1:\sempred{\_every\_q}(x)$, \\
$l_{41}:a_{41}:\sempred{\_fat\_a}(e')$,\\
$l_{42}:a_{42}:\sempred{\_cat\_n}(x_3)$\\
$l_5:a_5:\sempred{\_chased\_v}(e_{past})$, \\
$l_6:a_6:\sempred{\_some\_q}(x_4)$, \\
$l_9:a_9:\sempred{\_dog\_n}(x_5)$
\todo{add more stuff with ARG and equalities here}
\label{ex:cat-partial-parser}
\end{examples}

This \rmrs\ uses two new types of atoms.  Atoms of the form $x=y$
express that the two variables $x$ and $y$ should really be the same;
in the example, the \rmrs\ uses two variable names $x$ and $x'$, which
are mapped to the same variable $x$ in Fig.~\ref{fig:1}.  \todo{I'm
  making these names up -- adjust to RMRS.}  Atoms of the form
$\ARG_i(a,x)$ and $\ARG_i(a,h)$ express that a node representing the
variable $x$ or a node denoted by the \emph{hole} $h$ is the $i$-th
child of the node to which the anchor $a$ refers.  The separation of
EPs and $\ARG$ atoms is perhaps the most important difference between
\rmrs\ and earlier underspecification formalisms such as \mrs\ and
dominance graphs, which required that a predicate and its arguments
had to be specified together.  \rmrs\ goes beyond these formalisms in
expressive power, especially if we allow atoms of the form
$\ARG_{\{2,3\}}(a,x)$ to express uncertainty as to whether $x$ is the
second or third child of the anchor $a$.

Notice also that the labels of the EPs for ``fat'' and ``cat'' are
stipulated to be equal in the \rmrs, whereas the anchors are not.
In the tree, it is the anchors that are mapped to the nodes with the
labels $\sem{\_fat\_a\_1}$ and $\sem{\_cat\_n\_1}$; the label is
mapped to the conjunction node just above them.  In other words, the
role of the anchor in an EP is to carry the predicate and connect to
the arguments; the role of the label is to connect the EP to the
surrounding formula.  The use of equal labels for two EPs stems from
\mrs, where intersective modification was also modeled by using two
EPs with the same label.

Finally, here is an \rmrs\ that could be derived from a deep grammar
such as the ERG:

\begin{examples}
\item $l_1:a_1\handel\mbox{\_every\_q\_1}(x_1),$\\
\hspace*{0.1in}$\mbox{RSTR}(a_1,h_2),
\mbox{BODY}(a_1,h_3)$\\ 
$l_{41}:a_{41}\handel\mbox{\_fat\_j\_1}(e'), \mbox{ARG1}(a_{41},x_2)$\\
$l_{42}:a_{42}\handel\mbox{\_cat\_n\_1}(x_3)$\\
$l_5:a_5\handel\mbox{\_chase\_v\_1}(e_{spast})$,\\
\hspace*{0.1in}$\mbox{ARG1}(a_5,x_4),
\mbox{ARG2}(a_5,x_5)$\\ 
$l_6:a_6\handel\mbox{\_some\_q\_1}(x_6)$,\\
\hspace*{0.1in}$\mbox{RSTR}(a_6,h_7),
\mbox{BODY}(a_6,h_8)$\\ 
$l_9:a_9\handel\mbox{\_dog\_n\_1}(x_7)$\\
$h_2=_q l_{42}, l_{41}=l_{42}, h_7 =_q l_9$\\
$x_1=x_2, x_2=x_3, x_3=x_4,$\\
$x_5=x_6, x_5=x_7$
\todo{fix RSTR and BODY to ARG1 and ARG2? or talk about it in the text?}
\label{ex:cat-erg}
\end{examples}

This is now an extremely detailed underspecified representation of the
different semantic representations of the sentence, which is
essentially a notational variant of the \mrs\ that the ERG would
derive.  Notice the use of atoms like $h_2 \qeq l_{42}$, which specify
a kind of ``outscopes'' relationship between the hole and the label,
and can be used to underspecify the scope of the two quantifiers.

In summary, \rmrs\ is a formalism for representing partial information
about semantic representations.  It generalizes \mrs\ in that it can
represent the output of a deep grammar in a way that is a notational
variant of \mrs; but in cases in which a deep grammar is not
available, it can be used to capture whatever partial semantic
information a shallower NLP tool was able to provide.  \rmrs\ is
designed to serve as a generic formalism for partial semantic
representations which abstracts over the details of different NLP
tools and allows us to represent, process, and compare information
from different tools in the same formalism.  In the long run, it could
also be a platform for incrementally enriching partial information
such as (\ref{ex:cat-pos}) in order to eventually approximate more
detailed representations such as (\ref{ex:cat-partial-parser}).

\todo{Leidensdruck!}
